---
title: "P8105_hw3"
author: "Jingwei Ren"
date: "10/12/2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
```

##problem1
First, do some data cleaning:

format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Excellent” to “Poor”

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health" & (response == "Excellent" | response == "Very good" |
                                       response == "Good" | response == "Fair" | 
                                       response == "Poor")) %>% 
  mutate(response = factor(response, level = c("Excellent","Very good", "Good", "Fair", "Poor")))
```

Using this dataset, do or answer the following (commenting on the results of each):

In 2002, which states were observed at 7 locations?
```{r}
brfss %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  distinct(locationdesc) %>% 
  count() %>% 
  filter(n == 7) 
```
CT,FL and NC were observed at 7 locations.

Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.
Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
